---
title: "t-SNE plots for GTEX tissues"
author: "Stephens Lab"
date: "June 19, 2015"
output: html_document
---

```{r packages,echo=FALSE}
suppressMessages(suppressWarnings(library(tsne)))
suppressMessages(suppressWarnings(library(qtlcharts)))
suppressMessages(suppressWarnings(library(data.table)))
suppressMessages(suppressWarnings(library(maptpx)))

#library(iplots)

```


The t-SNE plot of the 8555 GTEX samples based on the reads data of the top 5000 genes based on the mean expression levels is given as follows 

```{r tissue_tSNE}

sample_id_sub <- read.table("../data/samples_id.txt")[,3];

males_identifier=as.numeric(as.matrix(read.table("../data/males_identifier.txt")));
females_identifier=as.numeric(as.matrix(read.table("../data/females_identifier.txt")));

gender_label=array(0,length(males_identifier)+length(females_identifier));

gender_label[males_identifier]="MALE";
gender_label[females_identifier]="FEMALE";
sample_id_sub_gender=paste0(sample_id_sub,"_",gender_label);

#reads.data=data.frame(fread("../data/reads_expression_top_5000_genes.txt"))[,-1];

#reads.data_100_genes=reads.data[,1:100];

tsne_samples <- read.table("../data/tsne_samples.txt");

options(warn=-1)
suppressWarnings(suppressMessages(iplot(tsne_samples[,1],tsne_samples[,2],as.numeric(sample_id_sub),sample_id_sub_gender)))


#iplotCurves(log(reads.data_100_genes+1), 1:100, tsne_samples,
 #           group=as.numeric(sample_id_sub));




```

The colors in the above plot are by the tissue labels.

We then fitted the Admixture model and applied the t-SNE on the topic proportion matrix of the Admixture model. It is presented below.

```{r topicmodel_tsne_tissue}


topicmodel_tsne_15 = read.table("../data/topicmodel_samples_tsne_15.txt");

suppressWarnings(suppressMessages(iplot(topicmodel_tsne_15[,1],topicmodel_tsne_15[,2],as.numeric(sample_id_sub),sample_id_sub_gender)))


```

The colors in the above plot are by the tissue labels.
