---
title: "lsr.tissue_specific_effects"
author: "StephensLab"
date: "June 19, 2015"
output: html_document
---

```{r packages,echo=FALSE}
suppressMessages(suppressWarnings(library(tsne)))
suppressMessages(suppressWarnings(library(qtlcharts)))
suppressMessages(suppressWarnings(library(data.table)))
#library(iplots)

```

## General plot of the t-scores 

We plot the eQTL t-scores across the 44 tissues for 3000 cis-eQTLs in the file.

```{r t-score_plot}

tscore_data=read.table("../data/complete_tscore_eQTL.table.txt");

iplotCurves(tscore_data[1:3000,-1],1:44);

```

We extract the lfspr values for each gene across the different tissues and then we look at the tissue specific effects (if any). Possoble choices of tissues that show tissue specific effects are thyroid, lymphocytes, whole blood and testis. We want to figure out the eQTLs which are specific only to these tissues.

```{r read_data}

lfsr_data=read.table("../data/lfsr.et_withbma.txt");


post.mean.shrink = read.table("../data/post.mean.et_withbma.normalized.txt");

##  read the tissue names

names_tissues <-  read.table("../data/names_tissues.txt")[-1];

tissue_names_filtered=unlist(lapply(as.matrix(names_tissues),
                            function(x) substring(x,32,nchar(x)-7)));

gene_info_eqtl = data.frame(fread("../data/gene_info_eqtl.txt"));

genesnp.names=as.vector(as.matrix(read.table("../data/gene_snp_names.txt")));
gene_names_GTEX=substring(genesnp.names,1,15);

matching_labels_eqtls_to_gene_info= match(gene_names_GTEX,gene_info_eqtl[,1]);


index_not_found=which(is.na(matching_labels_eqtls_to_gene_info)==TRUE);


matching_labels=matching_labels_eqtls_to_gene_info[which(!is.na(matching_labels_eqtls_to_gene_info))];

flag=paste0(gene_info_eqtl[matching_labels,1],", ",
                      "name-",gene_info_eqtl$Associated.Gene.Name[matching_labels],", ",
                      "descr-",gene_info_eqtl$Description[matching_labels],",",
                      "chr-",gene_info_eqtl$Chromosome.Name[matching_labels]);

post.mean.shrink.matched=post.mean.shrink[-index_not_found,];
rownames(post.mean.shrink.matched)=flag;


lfsr_data.matched=lfsr_data[-index_not_found,];


```

## Lymphocytes specific effects

```{r lymphocytes_strong}

index_tissue=which(tissue_names_filtered=="Cells_EBV-transformed_lymphocytes");


thresh=0.05

  ##create a matrix showing whether or no lfsr satisfies threshold

sigmat =lfsr_data.matched <= thresh;


sigs=which(rowSums(sigmat)==1)

sigs.it=which(lfsr_data.matched[sigs,index_tissue]<thresh)

iplotCurves(post.mean.shrink.matched[sigs[sigs.it],],1:44)
```

## Thyroid Specific effects

```{r thyroid_strong}

index_tissue=which(tissue_names_filtered=="Thyroid");


thresh=0.05

  ##create a matrix showing whether or no lfsr satisfies threshold

sigmat =lfsr_data.matched <= thresh;


sigs=which(rowSums(sigmat)==1)

sigs.it=which(lfsr_data.matched[sigs,index_tissue]<thresh)

iplotCurves(post.mean.shrink.matched[sigs[sigs.it],],1:44)


```


## Whole Blood Specific effects

```{r whole_blood_strong}

index_tissue=which(tissue_names_filtered=="Whole_Blood");


thresh=0.05

  ##create a matrix showing whether or no lfsr satisfies threshold

sigmat =lfsr_data.matched <= thresh;


sigs=which(rowSums(sigmat)==1)

sigs.it=which(lfsr_data.matched[sigs,index_tissue]<thresh)


iplotCurves(post.mean.shrink.matched[sigs[sigs.it],],1:44)

```


## Testis strong effects

```{r testis_strong}

index_tissue=which(tissue_names_filtered=="Testis");


thresh=0.05

  ##create a matrix showing whether or no lfsr satisfies threshold

sigmat =lfsr_data.matched <= thresh;


sigs=which(rowSums(sigmat)==1)

sigs.it=which(lfsr_data.matched[sigs,index_tissue]<thresh)


iplotCurves(post.mean.shrink.matched[sigs[sigs.it],],1:44)


```


## Brain strong effects

```{r brain_strong}


index_tissue=7:16;


thresh=0.05

  ##create a matrix showing whether or no lfsr satisfies threshold

sigmat =lfsr_data.matched <= thresh;

lfsr_red=t(apply(lfsr_data.matched,1,function(x){c(x[1:6],min(x[7:16]),x[17:44])}))


sigs=which(rowSums(sigmat)==1)

sigs.it=which(lfsr_red[sigs,index_tissue]<thresh)


iplotCurves(post.mean.shrink.matched[sigs[sigs.it],],1:44)



```


## Tissue consistent effects

```{r no_sig}


index_tissue=1:44;


thresh=0.05

  ##create a matrix showing whether or no lfsr satisfies threshold

sigmat =lfsr_data.matched <= thresh;


sigs=which(rowSums(sigmat)==44);


iplotCurves(post.mean.shrink.matched[sigs,],1:44)


```
