---
title: "plot.by.post.weights"
author: "StephensLab"
date: "June 19, 2015"
output: html_document
---

```{r packages,echo=FALSE}
suppressMessages(suppressWarnings(library(qtlcharts)))

library("qtlcharts")

```

Here, I plot a histogram of the sum of the posterior weights for each $U_{k}$ across all omega. Recall the components:

$U_{k}=1$ Corresponds to the 44x44 Identity Matrix
$U_{k}=2$ Corresponds the 44x44 Empirical Covariance Matrix (T'T)
$U_{k}=2$ Corresponds to the rank 2 PCA Approximation
$U_{k}=3-8$ Corresponds to single rank SFA Approximation
$U_{k}=9$ Corresponds to rank 5 Approximation
$U_{k}=10-53$ Correspond to Singleton Configs
$U_{k}=54$ Corresponds to full configuration

The overarching theme is that the majority of SNPS find their maximal responsibility at one of the fuller rank components (i.e., XtX or rank 5 SFA ).
```{r read_data}

post.weights=readRDS("../data/ET_model_stronggenesnps/post.weight.et_withbma.rds")
genesnpnames=read.table("../data//ET_model_stronggenesnps/genesnpnames.txt")[,1]
tissue.names=read.table("../data//ET_model_stronggenesnps/tissuenames.txt")[,1]
post.mean.shrink = read.table("../gtex.Kushal/data/post.mean.et_withbma.normalized.txt");
rownames(post.mean.shrink)=genesnpnames
colnames(post.mean.shrink)=tissue.names

##Find the maximum responsibilitity for each componenet
weightsort=apply(post.weights,1,function(x){
  weightmat=matrix(x,ncol=54,nrow=22,byrow=TRUE)
  which.max(colSums(weightmat))
})


hist(weightsort,labels=seq(1:44),freq=FALSE,nclass=50,las=2,ylim=c(0,0.9),main="Maximum Responsibility over Omega")
text(x=2,y=0.6,"Uk=2")
text(x=9,y=0.4,"Uk=9")
text(x=27,y=0.2,"Uk=27")
text(x=49,y=0.1,"Uk=49")




#You can visit the interactive histogram at here##

# set_credentials_file("surbut", "wpbuudeooe")
# 
# py=plotly()
# library(plotly)
# 
# data <- list(
#   
#     x = weightsort,
#     type = "histogram"
#     
#   )
#   
#   layout <- list(
#   title = "Maximum Posterior Weight Across Omega",
#   xaxis = list(title = "UK"),
#   yaxis = list(title = "Count"),
#   barmode = "overlay",
#   bargap = 0.25,
#   bargroupgap = 0.3
# )
# response <- py$plotly(data, kwargs=list(filename="basic-histogram", fileopt="overwrite"))
# url <- response$url
# 
# 


```

Now, we might want to observe the pattern of effects (i.e., shrunken posterior means of t statistics) for SNPs with high responsibility at a particular component type. First, I plot eQTL with maximum loading on the empirical covariance matrix and on the rank 5 SFA approximation.

```{r,xtxers}
###Plot the ones with maximal loading on XtX
xtxers=which(weightsort==2)
empcov=sample(xtxers,1000)
iplotCurves(post.mean.shrink[empcov,],1:44)

#sfaers##
sfaers=which(weightsort==9)

iplotCurves(post.mean.shrink[sfaers,],1:44)
```

Now I plot those with maximum responsibility at U_k = 27, which corresponds to EBV (i.e, tissuenames 18).

```{r, EBV}
##27ers##
twentyseveners=which(weightsort==27)

iplotCurves(post.mean.shrink[twentyseveners,],1:44)
```

You can see that the other common loading is at $U_k$ = 44, which corresponds to testes.

```{r, TESTES}

fortyniners=which(weightsort==49)
iplotCurves(post.mean.shrink[fortyniners,],1:44)

```

