---
title: "tissue.specific"
output: html_document
---

First,load the data

```{r}
dirpath=("../data/ET_model_stronggenesnps/")
lfsr = read.table(paste0(dirpath,"lfsr.et_withbma.txt"))
post.means=read.table(paste0(dirpath,"post.mean.et_withbma.txt"))
se.gp.hat=na.omit(read.table(paste0(dirpath,"max_sebeta_eQTL.table.txt")))[,-c(1)]
s.j=se.gp.hat[,-1]#remove genesnpnames
t.stat=na.omit(read.table(paste0(dirpath,"max_tscore_eQTL.table.txt")))[,-c(1,47)]
t.j=t.stat[,-1]
marginal.var=read.table(paste0(dirpath,"marginal.var.et_withbma.txt"))

```

We might want to choose a few tissue specific examples to plot. I choose the example that showed the maximal LFSR (because most tissues were null). I plot the posterior mean as converted from the ET model and also plot the mle (bhat and tstat).

```{r, echo=FALSE}
plot.func.t=function(j,posterior.means,lfsr.mat,marginal.var,s.j,t.j){
genesnpnames=rownames(posterior.means)
tissue.names=colnames(posterior.means)
R=ncol(posterior.means)
posterior.means=as.matrix(posterior.means)

s.j=as.matrix(s.j)
  col.mat=NULL
  
  for(r in 1:R){
    
    if (lfsr.mat[j,r]<=0.10) {
      col.mat[r]=1
    } else if (lfsr.mat[j,r]<0.5) {
      col.mat[r]=2
    } else if (lfsr.mat[j,r]>=0.50) {
      col.mat[r]=3
    } 
  }
  


post.mean.converted=posterior.means[j,]*s.j[j,]
mar.sd.converted=s.j[j,]*sqrt(marginal.var[j,])

b=barplot(post.mean.converted,main=paste0("Posterior Tissue Means",genesnpnames[j]),names=tissue.names,col=col.mat,las=2,ylim=c((min(post.mean.converted)-0.05),(max(post.mean.converted)+0.10)),cex.names=0.5,ylab="Posterior Mean")

  mean=as.numeric(post.mean.converted)
  sd=as.numeric(mar.sd.converted)
  segments(b, mean - sd, b, mean + sd, lwd=2)
  t=as.numeric(levels(as.factor(col.mat)))
  key=c("<0.10","0.10<x<0.50",">0.5")
  legend("topright",legend=key[t],col=t,pch=1,title="lfsr")

mle=as.matrix(t.j[j,]*s.j[j,])
bhat=barplot(mle,main=paste0("b.hat",genesnpnames[j]),names=tissue.names,las=2,ylim=c((min(mle)-0.05),(max(mle)+0.10)),cex.names=0.5,ylab="MLE")
t=barplot(as.matrix(t.j[j,]),main=paste0("t.stat",genesnpnames[j]),names=tissue.names,las=2,ylim=c((min(t.j[j,])-0.05),(max(t.j[j,])+0.10)),cex.names=0.5,ylab="MLE")


}


plot.func.tneg=function(j,posterior.means,lfsr.mat,marginal.var,s.j,t.j){
genesnpnames=rownames(posterior.means)
tissue.names=colnames(posterior.means)
R=ncol(posterior.means)
posterior.means=as.matrix(posterior.means)

s.j=as.matrix(s.j)
  col.mat=NULL
  
  for(r in 1:R){
    
    if (lfsr.mat[j,r]<=0.10) {
      col.mat[r]=1
    } else if (lfsr.mat[j,r]<0.5) {
      col.mat[r]=2
    } else if (lfsr.mat[j,r]>=0.50) {
      col.mat[r]=3
    } 
  }
  


post.mean.converted=posterior.means[j,]*s.j[j,]
mar.sd.converted=s.j[j,]*sqrt(marginal.var[j,])

b=barplot(post.mean.converted,main=paste0("Posterior Tissue Means,",genesnpnames[j]),names=tissue.names,col=col.mat,las=2,ylim=c((min(post.mean.converted)-0.05),0),cex.names=0.5,ylab="Posterior Mean")

  mean=as.numeric(post.mean.converted)
  sd=as.numeric(mar.sd.converted)
  segments(b, mean - sd, b, mean + sd, lwd=2)
  t=as.numeric(levels(as.factor(col.mat)))
  key=c("<0.10","0.10<x<0.50",">0.5")
  legend("topright",legend=key[t],col=t,pch=1,title="lfsr")

mle=as.matrix(t.j[j,]*s.j[j,])
bhat=barplot(mle,main=paste0("b.hat",genesnpnames[j]),names=tissue.names,las=2,ylim=c((min(mle)-0.05),0),cex.names=0.5,ylab="MLE")
t=barplot(as.matrix(t.j[j,]),main=paste0("t.stat",genesnpnames[j]),names=tissue.names,las=2,ylim=c(min(t.j[j,]),0),cex.names=0.5,ylab="MLE")


}
```

Now, let's find some tissue specific snps.

```{r}
tissue.specific=which(lfsr==max(lfsr),arr.ind=TRUE)[1]


plot.func.t(tissue.specific,posterior.means = post.means,lfsr.mat = lfsr,marginal.var = marginal.var,s.j=s.j,t.j)

max.var=which.min(apply(lfsr,1,function(x){mean(x)}))
plot.func.tneg(max.var,posterior.means = post.means,lfsr.mat = lfsr,marginal.var = marginal.var,s.j=s.j,t.j)

